Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-11-13 files/Chapter3/dynamicconvergence.tex
chapter 3 fix

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
\textcolor{red}{1973 is die original, en 1997 is 'n update tot die 1973 met extra voorbeelde, kode en oefeninge. Die 1997 is die die kopie wat ek het. So prof moet se as ek 1973 eerder moet gebruik.}

In this chapter, two important convergence results of the Finite Element Method
(FEM) is discussed. The first result looks at the convergence of the Galerkin
approximation for second order hyperbolic type problems, as presented
in~\cite{BV13}. The second result, from the textbook~\cite{SF97}, examines the
convergence of eigenvalues and eigenfunctions for a one-dimensional vibration
problem, using the Finite Element Method.

\section{Convergence of the Galerkin approximation}\label{sec:convergence-of-the-galerkin-approximation}
In the article \cite{BV13}, the authors investigate the convergence of the
Galerkin approximation for second order hyperbolic type problems. The article
\cite{BSV17} extends the work of \cite{BV13} by including general damping and
damping at the endpoints. For the models in Chapter 1, \cite{BV13} is
sufficient while \cite{BSV17} provides more insight and improved notation.

@@

\subsubsection*{Problem GVar}
Given a function $f:J\rightarrow X$, find a function $u\in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to
$\Vert \cdot \Vert_{W}$ and for each $t\in J,\ u(t)\in V,\ u'(t) \in V,\ u''(t)\in W$ and
\begin{eqnarray}
	c(u''(t),v)+a(u'(t),v)+b(u(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in V, \label{GV_PB1}
\end{eqnarray}
while $u(0)=u_{0}$, $u'(0)=u_{1}$.

@@

This theorem is presented and proved in \cite{BV13}.

@@

Since the set $\{\phi_1, \ \phi_2,...,\phi_n\}$ forms a linearly independent
basis for $S^h$, the bilinear form $c$ is positive definite on $S^h$.
Consequently, the matrix $M$ is invertible. The equation \eqref{ODE} can
therefore be rewritten as $\bar{u}'' + M^{-1}C\bar{u}' + M^{-1}K \bar{u} =
	M^{-1}F(t)$. Using the theory of linear differential equations, a unique
solution for Problem ODE can be found, proving Proposition 1.

@@

Let $u$ be the solution of Problem GVar and $u^h$ be the solution of Problem
GVar$^h$. The authors of \cite{BV13} define the following error,

@@

\subsection{Main result}
\textcolor{red}{************** NIE DIE MAIN RESULT NIE*******************}

The main result proving the convergence of the solution of the Galerkin
Approximation is given in \cite{BV13} as follows.

\newtheorem*{DC_Thm5}{Theorem 2}
\begin{DC_Thm5}
	{Main Result}\\
	If $f\in C^{2}([0,T]4,X)$, then
	\begin{eqnarray*}
		\Vert u(t_{k})-u_{k}^{h}\Vert_{W}\ & \leq & 7T^{2}\tau^{2}\max\Vert u_{h}^{(4)}\Vert_{W}+7T\tau^{2}\max\Vert u_{h}'''\Vert_{W}\\
		& & + \sqrt{2K_{a}}\tau^{4}\max\Vert u_{h}'''\Vert_{W}
	\end{eqnarray*}
	for each $t\in(0,\ T)$.
\end{DC_Thm5}

The parameter $\tau$ is some number in $(0,T)$. The constant $K_\alpha$ is the
constant defined in assumption \textbf{A4} of Section 2.2.1.

\subsection{Fully discrete Galerkin finite element approximation}
\textcolor{red}{********************HOU VOORLOPIG**********************}

Consider the time interval $J = [0,T]$. Divide $J$ into $N$ steps with length
$\tau = \frac{T}{N}$. Each interval can be expressed as $[t_{k-1}, \ t_k]$ for
$k = 1,...,N$. Denote the approximation of $u_h$ on the interval $[t_{k-1}, \
			t_k]$ by $u_k^h$, i.e. $u_h(t_k) = u_k^h$ for each $k$.\\

The error $e(t)$, can now be expressed as
\begin{eqnarray}
	e(t_k) = u(t_k) - u^h_k & = & [u(t_k)-u_h(t_k)] + [u_h(t_k) - u^h_k]. \label{error_ek}
\end{eqnarray}

@@

\textcolor{red}{********************HOU VOORLOPIG**********************}

\textcolor{red}{Dupliseering maar moontlik beter as werk net voor Main Result, p52}

\subsubsection{Error estimate for the fully discrete problem}

\textcolor{red}{Moontlik hou!}\\
There error estimate for $[u_h(t_k) - u^h_k]$ can now be determined. Recall that $\tau = \frac{T}{N}$. Subsituting $t = t_k$ and $t = t_k+1$ in Problem GVar$^h$ results in
\begin{eqnarray*}
	c( \tau^{-1}[v_{h}(t_{k+1})-v_{h}(t_{k})], \psi)+\frac{1}{2}a([v_{h}(t_{k+1})+v_{h}(t_{k})], \psi)+\frac{1}{2}b([u_{h}(t_{k+1})+u_{h}(t_{k})], \psi) \\
	=\frac{1}{2}([f(t_{k+1})+f(t_{k})], \psi)_{X}+c(\rho_{k}, \psi)
\end{eqnarray*}
where $v_{h}(t)=u_{h}'(t)$ and
\begin{eqnarray*}
	\rho_{k}=\tau^{-1}[v_{h}(t_{k+1})-v_{h}(t_{k})]-\frac{1}{2}[v_{h}'(t_{k+1})+v_{h}'(t_{k})].
\end{eqnarray*}

The authors in \cite{BV13} denote the following errors
\begin{eqnarray*}
	e_{k}=u_{h}(t_{k})-u_{k} \ \ \ \ \textrm{ and } \ \ \ \ q_{k}=u_{h}'(t_{k})-v_{k}.
\end{eqnarray*} In the authors approach, $e_0$ and $q_0$ are zero.

\newtheorem*{DC_Lem3}{Lemma 2}
\begin{DC_Lem3}
	Stability
	\begin{eqnarray*}
		\max\Vert e_{n}\Vert_{W}^{2}\leq 8T\tau\sum_{n=0}^{N-1}\Vert\epsilon_{n}\Vert_{W}^{2}+2\tau^{4}\Vert\rho_{0}\Vert_{W}^{2}+(8\tau^{2}+2\tau^{4}K_{a})\Vert\sigma_{0}\Vert_{W}^{2}.
	\end{eqnarray*}
\end{DC_Lem3}

Using this lemma, \cite{BV13} proves the following error estimate.\\
\textcolor{red}{*************************************}

@@

The parameter $\tau$ is some number in $(0,T)$. The constant $K_\alpha$ is the constant defined in assumption \textbf{A4} of Section 2.2.1.

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}

In this chapter, two important convergence results of the Finite Element Method
(FEM) is discussed. The first result looks at the convergence of the Galerkin
approximation for second order hyperbolic type problems (or general vibation problems). The papers considered are \cite{BV13} and \cite{BSV17}. The second result is from the textbook~\cite{SF73} and examines the
convergence of eigenvalues and eigenfunctions for a one-dimensional vibration
problem, using the Finite Element Method.

\section{Galerkin approximation for second order hyperbolic type problems}\label{sec:convergence-of-the-galerkin-approximation}
In the article \cite{BV13}, the authors investigate the convergence of the
Galerkin approximation for second order hyperbolic type problems. The article
\cite{BSV17} extends the work of \cite{BV13} by including general damping and
damping at the endpoints. For the models in Chapter 1, \cite{BV13} is
sufficient while \cite{BSV17} provides more insight and improved notation.

@@

\subsection{Formulation of the Galerkin approximation}
\subsubsection*{Problem GVar}
Given a function $f:J\rightarrow X$, find a function $u\in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to
$\Vert \cdot \Vert_{W}$ and for each $t\in J,\ u(t)\in V,\ u'(t) \in V,\ u''(t)\in W$ and
\begin{eqnarray}
	c(u''(t),v)+a(u'(t),v)+b(u(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in V, \label{GV_PB1}
\end{eqnarray}
while $u(0)=u_{0}$, $u'(0)=u_{1}$.

@@

@@

@@

@@

\textbf{Fully discrete Galerkin finite element approximation}

@@

A stability result is needed.

\newtheorem*{DC_Lem3}{Lemma}
\begin{DC_Lem3}
	Stability
	\begin{eqnarray*}
		\max\Vert e_{n}\Vert_{W}^{2}\leq KT\tau
	\end{eqnarray*}
	where $K$ depends on $u$, $u_h$ and their derivatives.
\end{DC_Lem3}

Using this lemma, \cite{BV13} prove the finial error estimate.

@@

\subsection{Main result}
The main result proving the convergence of the solution of the Galerkin
Approximation is given in \cite{BV13} as follows.
\newtheorem*{DC_Thm5}{Theorem 2}
\begin{DC_Thm5}
	{Main Result}\\
	If $f\in C^{2}([0,T],X)$, then
	\begin{eqnarray*}
		\Vert u(t_{k})-u_{k}^{h}\Vert_{W}\ & \leq & K\tau^2
	\end{eqnarray*}
	for each $t\in(0,\ T)$.
\end{DC_Thm5}

The constant $K$ depend on $u_h$ and its derivatives.

### 2023-11-13 files/Chapter3/femconvergence.tex
appendix
chapter 3 fix

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Convergence of eigenvalues and eigenfunctions}
In this section, the work in the textbook \cite[p.228-236]{SF73} is discussed. The results are the same as given in the textbook, how ever the proofs are expanded for clarity.\\

@@

The following eigenvalue is given in \cite{SF73}. It is important to note that the problem is limited to one-dimensional problems.

@@

\subsection{Preliminary theory and notation}
Some theory is required before the main results can be proven. The theory is from \cite{SF73}.\\

@@

From the textbook \cite{ORXX}, the following approximation theorem for finite elements is given verbatim:
\newtheorem*{Interpolation}{Theorem}
\begin{Interpolation}[The Interpolation Theorem for Finite Elements] 
	Let $\Omega$ be an open bounded domain in $\mathcal{R}^n$ satisfying the cone condition. Let $k$ be a fixed integer and $m$ an integer such that $0\leq m \leq k+1$. Let $\Pi \in L(H^{k+1}(\Omega), H^{m}(\Omega))$ be such that
	\begin{eqnarray}
	    \Pi u = u \ \ \ \ \textrm{ for all } u \in \mathcal{P}_k(\Omega)
	\end{eqnarray}
	Then for any $u \in H^{k+1}(\Omega)$ and for sufficiently small $h$, there exists positive a constant $C$, independent of $u$ and $h$, such that
	\begin{eqnarray}
	    ||u - \Pi u||_{H^m(\Omega)} \leq C \frac{h^{k+1}}{p^m} |u|_H^{k+1}(\Omega)
	\end{eqnarray}
	where $|u|_H^{k+1}(\Omega)$ is the seminorm.
\end{Interpolation}

For the requirements of this dissertation, this can be simplified with assumptions. The first assumption is a reiteration that the eigenvalue problem is one-dimensional. Another assumption is that the basis of $S^h$ consists of polynomials. With these assumptions, the semi-norm $|u|_H$ is equal to the norm $||u||$. This approximation theorem can be rewritten as
\newtheorem{Interpolation_2}{Theorem}
\begin{Interpolation_2}
	Suppose there exists an integer $k$ such that for each element
	\begin{equation*}
		s(\Pi_e) + 1 \leq k \leq r(\Pi_e) +1.
	\end{equation*}
	Then there exists a constant $C$ such that for any $u \in C^k_+(I)$,
	\begin{equation*}
		||(\Pi u)^{(m)} - u^{(m)}|| \leq C h^{k-m} || u^{(k)}|| \ \text{ for } m = 0,1,...,k.
	\end{equation*}
\end{Interpolation_2}

This will be referred to as the approximation theorem in this section.\\

@@

Since $S^h$ is finite dimensional, the ordering of the eigenvalues still hold. The eigenvalues in the space $S^h$ is denoted as $\lambda_i^h$ with corresponding normalized eigenfunction $u^h_i$ for $i = 1,2,...,m$.\\

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{FEM comparison of eigenvalues and eigenfunctions}

@@

The following eigenvalue is given in \cite{SF73}.

@@

Galerkin approximation for \eqref{GEVP}:
For $u^h \in S^h$ and $b(u^h, v) = \lambda^h(u^h,v)$ for all $v \in S^h$. For detail, see Chapter 5.

\subsection{Preliminary theory and notation}
In this section, the work in the textbook \cite[p.228-236]{SF73} is discussed. The results are the same as given in the textbook, how ever the proofs are expanded for clarity.

Some theory is required before the main results can be proven. The theory is from \cite{SF73}.\\

@@

From the textbook \cite{OR76}, the following approximation theorem for finite elements is given verbatim:
\newtheorem*{Interpolation}{Theorem}
\begin{Interpolation}[The Interpolation Theorem for Finite Elements] 
	Let $\Omega$ be an open bounded domain in $\mathcal{R}^n$ satisfying the cone condition. Let $k$ be a fixed integer and $m$ an integer such that $0\leq m \leq k+1$. Let $\Pi \in L(H^{k+1}(\Omega), H^{m}(\Omega))$ be such that
	\begin{eqnarray}
	    \Pi u = u \ \ \ \ \textrm{ for all } u \in \mathcal{P}_k(\Omega)
	\end{eqnarray}
	Then for any $u \in H^{k+1}(\Omega)$ and for sufficiently small $h$, there exists positive a constant $C$, independent of $u$ and $h$, such that
	\begin{eqnarray}
	    ||u - \Pi u||_{H^m(\Omega)} \leq C \frac{h^{k+1}}{p^m} |u|_H^{k+1}(\Omega)
	\end{eqnarray}
	where $|u|_H^{k+1}(\Omega)$ is the seminorm.
\end{Interpolation}

For the requirements of this dissertation, this can be simplified with assumptions. Another assumption is that the basis of $S^h$ consists of polynomials. With these assumptions, the semi-norm $|u|_H$ is equal to the norm $||u||$. This approximation theorem can be rewritten as
\newtheorem{Interpolation_2}{Theorem}
\begin{Interpolation_2}
	Suppose there exists an integer $k$ such that for each element
	\begin{equation*}
		s(\Pi_e) + 1 \leq k \leq r(\Pi_e) +1.
	\end{equation*}
	Then there exists a constant $C$ such that for any $u \in C^k_+(I)$,
	\begin{equation*}
		||(\Pi u)^{(m)} - u^{(m)}|| \leq C h^{k-m} || u^{(k)}|| \ \text{ for } m = 0,1,...,k.
	\end{equation*}
\end{Interpolation_2}

This will be referred to as the approximation theorem in this section.

@@

Since $S^h$ is finite dimensional, the ordering of the eigenvalues remains valid. The eigenvalues in the space $S^h$ is denoted as $\lambda_i^h$ with corresponding normalized eigenfunction $u^h_i$ for $i = 1,2,...,m$.\\

### 2023-11-13 files/Overview.tex
chapter 3 fix

\subsubsection{Chapter 3}
In this chapter, some theory for the Finite Element Method is discussed. This chapter contains two parts, that covers two different theoretical results. The first is on the convergence of the Galerkin approximation of a general vibration problem. This theory is presented in the article \cite{BV13}. The second part concerns the convergence of the eigenvalues and eigenfunctions of a vibration problem when applying the Finite Element Method. This theory is presented in the textbook \cite{SF97}\textcolor{red}{Nie goed genoeg vir dinamiese probleem}.

@@

The second part of the chapter considers work done in a textbook \cite{SF97} on eigenvalue problems for elliptic partial differential equations. The specific work discussed, covers the convergence of the eigenvalues and eigenfunctions of a one-dimensional vibration problem when applying the Finite Element Method. The authors consider a general eigenvalue problem. Then using the Rayleigh-quotient, from the Rayleigh-Ritz method, as well as an approximation theorem from \cite{OR97}, the main result is proven. The specific work done in this section is updating the notation of the textbook, as well as expanding some proofs so that the results are easier to understand.

@@

The next section is a short discussion on a suspended beam model, from an important article \cite{SP06}. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a physical beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The authors report on an experiment with a vibrated a free-free beam measure the natural frequencies. These emperical results are then compared to the theoretical results. This result, together with the results in chapter 6, give a good picture to the validity of the Timoshenko beam theory. Finally the chapter is concluded by looking at a model of a cantilever beam with a tip-body with elastic interfaces. These results combine modal analysis and investigates at the behavior of the model. This section is an example of the usefullness of the Timoshenko beam theory. A similar model using realistic two- or three-dimensional model will be very challenging.

@@

@@

@@

The second result concerns the convergence of the eigenvalues and eigenfunctions with the Finite Element Method for one-dimensional vibration problems. This is a result form the textbook \cite{SF97}.**********

@@

The second result concerns the convergence of the eigenvalue and eigenfunctions of a one-dimensional second order hyperbolic type problem when applying the Finite Element Method. This is presented in the textbook \cite{SF97}.

>>>>

\subsubsection{Chapter 3}
In this chapter, some theory for the Finite Element Method is discussed. This chapter contains two parts, that covers two different theoretical results. The first is on the convergence of the Galerkin approximation of a general vibration problem. This theory is presented in the article \cite{BV13}. The second part concerns the convergence of the eigenvalues and eigenfunctions of a vibration problem when applying the Finite Element Method. This theory is presented in the textbook \cite{SF73}\textcolor{red}{Nie goed genoeg vir dinamiese probleem}.

@@

The second part of the chapter considers work done in a textbook \cite{SF73} on eigenvalue problems for elliptic partial differential equations. The specific work discussed, covers the convergence of the eigenvalues and eigenfunctions of a general vibration problem when applying the Finite Element Method. The authors consider a general eigenvalue problem. Then using the Rayleigh-quotient, from the Rayleigh-Ritz method, as well as an approximation theorem from \cite{OR97}, the main result is proven. The specific work done in this section is updating the notation of the textbook, as well as expanding some proofs so that the results are easier to understand.

@@

\textcolor{red}{*************Copy van chapter 6******************}

\textcolor{red}{*******************************}

The next section is a discussion on the article \cite{SP06}. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a physical beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The authors report on an experiment with forced vibration a free-free beam where the natural frequencies are measured. These emperical results are then compared to the theoretical results. This result, together with the results in chapter 6, give a good picture to the validity of the Timoshenko beam theory. Finally the chapter is concluded by looking at a model of a cantilever beam with a tip-body with elastic interfaces. These results combine modal analysis and investigates at the behavior of the model. A similar model using realistic two- or three-dimensional model will be very challenging.

@@

\textcolor{red}{*************hou voorlopig******************}

@@

\textcolor{red}{*******************************}

@@

The second result concerns the convergence of the eigenvalues and eigenfunctions with the Finite Element Method for one-dimensional vibration problems. This is a result form the textbook \cite{SF73}.**********

@@

The second result concerns the convergence of the eigenvalue and eigenfunctions of a one-dimensional second order hyperbolic type problem when applying the Finite Element Method. This is presented in the textbook \cite{SF73}.
